---
title: Intro to Frequentist (Multilevel) Generalised Linear Models (GLM) in R with glm and lme4
author: '[Qixiang Fang](https://www.uu.nl/staff/QFang) and [Rens van de Schoot](https://www.rensvandeschoot.com/about-rens/)'
date: 'Last modified: date: `r format(Sys.time(), "%d %B %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
  
This tutorial provides the reader with a basic introduction to genearlised linear models (GLM) using the frequentist approach. Specifically, this tutorial focuses on the use of logistic regression in both binary-outcome and count/porportion-outcome scenarios, and the respective approaches to model evaluation. The tutorial uses the **Thai Educational Data** example in Chapter 6 of the book [Multilevel analysis: Techniques and applications](https://www.rensvandeschoot.com/multilevel-book/). Furthermore, the tutorial briefly demonstrates the multilevel extension of GLM models with the `lme4` package in R. Lastly, more distributions and link functions in the GLM framework are discussed.


This tutorial follows this structure:  
1. Preparation;  
2. Introduction to GLM;  
3. Thai Educational Data;  
4. Data Preparation;   
5. Binary (Bernoulli) Logistic Regression;   
6. Binomial Logistic Regression;   
7. Multilevel Logistic Regression;   
8. Other Family and Link Functions.  


Note that this tutorial is meant for beginners and therefore does not delve into technical details and complex models. For a detailed introduction into frequentist multilevel models, see this [LME4 Tutorial](https://www.rensvandeschoot.com/tutorials/lme4). For an extensive overview of GLM models, see [here](https://en.wikipedia.org/wiki/Generalized_linear_model). The Bayesian version of this tutorial can also be found [here](https://www.rensvandeschoot.com/generalised-linear-models-with-brms/).

This tutorial expects:  
- Installation of R packages [`lme4`](https://cran.r-project.org/web/packages/lme4/index.html) for multilevel modelling (this tutorial uses version 1.1-18-1);  
- Installation of R package [`tidyverse`](https://cran.r-project.org/web/packages/tidyverse/index.html) for data manipulation and plotting with ggplot2;  
- Installation of R package [`haven`](https://cran.r-project.org/web/packages/tidyverse/index.html) for reading sav format data;    
- Installation of R package [`jtools`](https://cran.r-project.org/web/packages/jtools/index.html) for handling of model summaries;  
- Installation of R package [`ggstance`](https://cran.r-project.org/web/packages/ggstance/index.html) for visualisation purposes;  
- Installation of R package [`ROCR`](https://cran.r-project.org/web/packages/ROCR/index.html) for calculating area under the curve (AUC);   
- Installation of R package [`sjstats`](https://cran.r-project.org/web/packages/sjstats/index.html) for calculating intra-class correlation (ICC). Remember to install version 0.17.5 (using the command `install_version("sjstats", version = "0.17.5")` after loading the package `devtools`, because the latest version of `sjstats` does not support the ICC function anymore);  
- Installation of R package [`effects`](https://cran.r-project.org/web/packages/effects/index.html) for plotting parameter effects;  
- Basic knowledge of hypothesis testing and statistical inference;  
- Basic knowledge of correlation and regression;  
- Basic knowledge of coding in R;  
- Basic knowledge of plotting and data manipulation with tidyverse.  

## **2. Introduction to Genearlised Linear Models (GLM)**
If you are already familar with generalised linear models (GLM), you can proceed to the next section. Otherwise, click "Read More" to learn about GLM. 

[expand title="Read More" trigclass="noarrow my_button" targclass="my_content" tag="button"]
  

Recall that in a linear regression model, the object is to model the expected value of a continuous variable, $Y$, as a linear function of the predictor, $\eta = X\beta$. The model structure is thus: $E(Y) = X\beta + e$, where $e$ refers to the residual error term. The linear regression model assumes that $Y$ is continuous and comes from a normal distribution, that $e$ is normally distributed and that the relationship between the linear predictor $\eta$ and the expected outcome $E(Y)$ is strictly linear. However, these assumptions are easily violated in many real world data examples, such as those with binary or proportional outcome variables and those with non-linear relationships between the predictors and the outcome variable. In these scenarios where linear regression models are clearly inappropriate, generalised linear models (GLM) are needed.


The GLM is the genearlised version of linear regression that allows for deviations from the assumptions underlying linear regression. The GLM generalises linear regression by assuming the dependent variable $Y$ to be generated from any particular distribution in an exponential family (a large class of probability distributions that includes the normal, binomial, Poisson and gamma distributions, among others). In this way, the distribution of $Y$ does not necessarily have to be normal. In addition, the GLM allows the linear predictor $\eta$ to be connected to the expected value of the outcome variable, $E(Y)$, via a link function $g(.)$. The outcome variable, $Y$, therefore, depends on $\eta$ through $E(Y) = g^{-1}(\eta) = g^{-1}(X\beta)$. In this way, the model does not assume a linear relationship between $E(Y)$ and $\eta$; instead, the model assumes a linear relationship between $E(Y)$ and the transformed $g^{-1}(\eta)$. 


This tutorial focuses on the probably most popular example of GLM: **logistic regression**. Logistic regression has two variants, the well-known **binary logistic regression** that is used to model binary outcomes (1 or 0; "yes" or "no"), and the less-known **binomial logistic regression** suited to model count/proportion data. 


Binary logistic regression assumes that $Y$ comes from a Bernoulli distribution, where $Y$ only takes a value of 1 (target event) or 0 (non-target event). Binary logistic regression connects $E(Y)$ and $\eta$ via the logit link $\eta = logit(\pi) = log(\pi/(1-\pi))$, where $\pi$ refers to the probability of the target event ($Y = 1$).

Binomial logistic regression, in contrast, assumes a binomial distribution underlying $Y$, where $Y$ is interpreted as the number of target events, can take on any non-negative integer value and is binomially distributed with regards to $n$ number of trials and $\pi$ probability of the target event. The link function is the same as that of binary logistic regression. 


[/expand]

  <p>&nbsp;</p>

The next section details the exampler data (**Thai Educational Data**) in this tutorial, followed by the demonstration of the use of both binary and binomial logistic regression. 


## **3. Thai Educational Data**
The data used in this tutorial is the Thai Eduational Data that is also used as an example in Chapter 6 of [Multilevel analysis: Techniques and applications](https://www.rensvandeschoot.com/multilevel-book/). The data can be downloaded from [here](https://github.com/MultiLevelAnalysis/Datasets-third-edition-Multilevel-book/blob/master/chapter%206/Thaieduc/thaieduc.sav). 


The data stems from a national survey of primary education in Thailand (Raudenbush & Bhumirat, 1992). Each row in the data refers to a pupil. The outcome variable `REPEAT` is a dichotomous variable indicating whether a pupil has repeated a grade during primary education. The `SCHOOLID` variable indicates the school of a pupil. The person-level predictors include: `SEX` (0 = female, 1 = male) and `PPED` (having had preschool education, 0 = no, 1 = yes). The school-level is `MSESC`, representing school mean SES (socio-economic status) scores.


The main research questions that this tutorial seeks to answer using the Thai Educational Data are: 


1. Ignoring the clustering structure of the data, what are the effects of gender and preschool education on whether a pupil repeats a grade?  
2. Ignoring the clustering structure of the data, what is the effect of school mean SES on the proportion of pupil repeating a grade?  
3. Considering the clustering structure of the data, what are the effects of gender, preschool education and school mean SES on whether a pupil repeats a grade?


These three questions are answered by using these following models, respectively: binary logistic regressioin; binomial logistic regression; multilevel binary logistic regression. 

## **4. Data Preparation**
#### 4.1. Load necessary packages
```{r, results='hide', message=FALSE, warning=FALSE}
# if you dont have these packages installed yet, please use the install.packages("package_name") command.
library(lme4) # for multilevel models
library(tidyverse) # for data manipulation and plots
library(haven) #for reading sav data
library(sjstats) #for calculating intra-class correlation (ICC)
library(effects) #for plotting parameter effects
library(jtools) #for transformaing model summaries
library(ROCR) #for calculating area under the curve (AUC) statistics
```

#### 4.2. Import Data
```{r}
ThaiEdu_Raw <- read_sav("https://github.com/MultiLevelAnalysis/Datasets-third-edition-Multilevel-book/blob/master/chapter%206/Thaieduc/thaieduc.sav?raw=true")
head(ThaiEdu_Raw)
```
Alternatively, you can download the data directly from [here](https://github.com/MultiLevelAnalysis/Datasets-third-edition-Multilevel-book/tree/master/chapter%206/Thaieduc) and import it locally. 


#### 4.3. Data Processing
```{r}
ThaiEdu_New <- ThaiEdu_Raw %>%
  mutate(SCHOOLID = factor(SCHOOLID),
         SEX = if_else(SEX == 0, "girl", "boy"),
         SEX = factor(SEX, levels = c("girl", "boy")),
         PPED = if_else(PPED == 0, "no", "yes"),
         PPED = factor(PPED, levels = c("no", "yes")))

head(ThaiEdu_New)
```


#### 4.4. Inspect Missing Data
```{r}
ThaiEdu_New %>%
  summarise_each(list(~sum(is.na(.)))) %>%
  gather()
```
The data has 1066 observations missing for the `MSESC` variable. The treatment of missing data is a complicated topic on its own. For the sake of convenience, we simply list-wise delete the cases with missing data in this tutorial.

```{r}
ThaiEdu_New <- ThaiEdu_New %>%
  filter(!is.na(MSESC))
```


## **5. Binary Logistic Regression**
#### 5.1. Explore Data: number of `REPEAT` by `SEX` and `PPED`
```{r}
ThaiEdu_New %>%
  group_by(SEX) %>%
  summarise(REPEAT = sum(REPEAT))
```

```{r}
ThaiEdu_New %>%
  group_by(PPED) %>%
  summarise(REPEAT = sum(REPEAT))
```

It seems that the number of pupils who repeated a grade differs quite a bit between the two genders, with more male pupils having to repeat a grade. More pupils who did not have preschool education repeated a grade. This observation suggests that `SEX` and `PPED` might be predictive of `REPEAT`.

#### 5.2. Fit a Binary Logistic Regression Model

`R` has the `base` package installed by default, which includes the `glm` function that runs GLM. The arguments for `glm` are similar to those for `lm`: `formula` and `data`. However, `glm` requires an additional argument: `family`, which specifies the assumed distribution of the outcome variable; within `family` we also need to specify the link function. The default of `family` is `gaussian(link = "identity")`, which leads to a linear model that is equivalent to a model specified by `lm`. In the case of binary logistic regression, `glm` requires that we specify a binomial distribution with the logit link, namely `family = binomial(link = "logit")`. 

```{r}
Model_Binary <- glm(formula = REPEAT ~ SEX + PPED,
                    family = binomial(link = "logit"),
                    data = ThaiEdu_New)
summary(Model_Binary)
```

#### 5.3. Interpretation
From the `summary` output above, we can see that `SEX` positively and significantly predicts a pupil's probability of repeating a grade, while `PPED` negatively and significantly so. Specifically, in comparison to being a girl, being a boy is more likely to repeat a grade. Having previous schooling is less likely to result in repeating a grade.



To interpret the value of the parameter estimates, we need to exponentiate the estimates. The `summ` function from the `jtools` packages provides an easy to do so for any model fitted by `glm`. See below.
```{r}
summ(Model_Binary, exp = T) # set "exp = T" to show esponentiated estimates; if you need standardised estimaets, set "scale = T"
```

Note that the interpretation of the parameter estimates is linked to the odds rather than probabilities. The definition of odds is: P(event occurring)/P(event not occurring). In this analysis, assuming everything else stays the same, being a boy increases the odds of repeating a grade by 54%, in comparison to being a girl; having preschool education lowers the odds of repeating a grade by (1 - 0.54)% = 46%, in comparison to not having preschool education, assuming everything else stays constant.

#### 5.4. Visualisation of Parameter Effects
To make the interpretation of the parameter effects even easier, we can use the `allEffects` function from the `effects` package to visualise the parameter effects. See below.
```{r}
plot(allEffects(Model_Binary))
```

Note that in both plots, the y scale refers to the probability of repeating a grade rather than the odds. Probabilities are more interpretable than odds. The probability scores for each variable are calculated by assuming that the other variables in the model are constant and take on their average values. As we can see, assuming that a pupil has an average preschool education, being a boy has a higher probability (~0.16) of repeating a grade than being a girl ~0.11). Likewise, assuming that a pupil has an average gender, having preschool education has a lower probability (~0.11) of repeating a grade than not having preschool education (~0.18). Note that in both plots the confidence intervals for the estimates are also included to give us some idea of the uncertainties of the estimates.


Note that the notion of average `preschool education` and `gender` may sound strange, given they are categorical variables (i.e. factors). If you are not comfortable with the idea of assuming an average factor, you can specify your intended factor level as the reference point, by using the `fixed.predictors = list(given.values = ...)` argument in the `allEffects` function. See below:

```{r}
plot(allEffects(Model_Binary, fixed.predictors = list(given.values=c(SEXboy=0, PPEDyes = 0))))
```

Setting `SEXboy = 0` means that for the PPED effect plot, the reference level of the `SEX` variable is set to 0; `PPEDyes = 0` results in the 0 being the reference level of the `PPED` variable in the SEX effect plot. 

Therefore, as the two plots above show, assuming that a pupil has no preschool education, being a boy has a higher probability (~0.20) of repeating a grade than being a girl ~0.14); assuming that a pupil is female, having preschool education has a lower probability (~0.09) of repeating a grade than not having preschool education (~0.15).

#### 5.5. Model Evaluation: Goodness of Fit

There are different ways to evaluate the goodness of fit of a logistic regression model.

##### 5.5.1. Likelihood ratio test

A logistic regression model has a better fit to the data if the model, compared with a model with fewer predictors, demonstrates an improvement in the fit. This is performed using the *likelihood ratio test*, which compares the likelihood of the data under the full model against the likelihood of the data under a model with fewer predictors. Removing predictor variables from a model will almost always make the model fit less well (i.e. a model will have a lower log likelihood), but it is useful to test whether the observed difference in model fit is statistically significant.

```{r}
#specify a model with only the `SEX` variable
Model_Binary_Test <- glm(formula = REPEAT ~ SEX,
                         family = binomial(link = "logit"),
                         data = ThaiEdu_New)

#use the `anova()` function to run the likelihood ratio test
anova(Model_Binary_Test, Model_Binary, test ="Chisq")
```

As we can see, the model with both `SEX` and `PPED` predictors provide a significantly better fit to the data than does the model with only the `SEX` variable. Note that this method can also be used to determine whether it is necessary to include one or a group of variables.

##### 5.5.2. AIC
The Akaike information criterion (AIC) is another measure for model selection. Different from the likelihood ratio test, the calculation of AIC not only regards the goodness of fit of a model, but also takes into account the simplicity of the model. In this way, AIC deals with the trade-off between goodness of fit and complexity of the model, and as a result, disencourages overfitting. A smaller AIC is preferred.
```{r}
Model_Binary_Test$aic
Model_Binary$aic
```
With a smaller AIC value, the model with both `SEX` and `PPED` predictors is preferred to the one with just the `SEX` predictor.

##### 5.5.3. Correct Classification Rate
The percentage of correct classification is another useful measure to see how well the model fits the data.
```{r}
#use the `predict()` function to calculate the predicted probabilities of pupils in the original data from the fitted model
Pred <- predict(Model_Binary, type = "response")
Pred <- if_else(Pred > 0.5, 1, 0)
ConfusionMatrix <- table(Pred, pull(ThaiEdu_New, REPEAT)) #`pull` results in a vector
#correct classification rate
sum(diag(ConfusionMatrix))/sum(ConfusionMatrix)

```

```{r}
ConfusionMatrix
```

We can see that the model correctly classifies 85.8% of all the observations. However, a closer look reveals that the model predicts all of the observations to belong to class "0", meaning that all pupils are predicted not to repeat a grade. Given that the majority category of the `REPEAT` variable is 0 (No), the model does not perform better in classification than simply assigning all observations to the majority class 0 (No). 

##### 5.5.4. AUC (area under the curve). 
An alternative to using correct classification rate is the Area under the Curve (AUC) measure. The AUC measures discrimination, that is, the ability of the test to correctly classify those with and without the target response. In the current data, the target response is repeating a grade. We randomly pick one pupil from the "repeating a grade" group and one from the "not repeating a grade" group. The pupil with the higher predicted probability should be the one from the "repeating a grade" group. The AUC is the percentage of randomly drawn pairs for which this is true. This procedure sets AUC apart from the correct classification rate because the AUC is not dependent on the imblance of the proportions of classes in the outcome variable. A value of 0.50 means that the model does not classify better than chance. A good model should have an AUC score much higher than 0.50 (preferably higher than 0.80).

```{r}
# Compute AUC for predicting Class with the model
Prob <- predict(Model_Binary, type="response")
Pred <- prediction(Prob, as.vector(pull(ThaiEdu_New, REPEAT)))
AUC <- performance(Pred, measure = "auc")
AUC <- AUC@y.values[[1]]
AUC
```

With an AUC score of 0.60, the model does not discriminate well.


## **6. Binomial Logistic Regression**
As mentioned in the beginning, logistic regression can also be used to model count or proportion data. Binary logistic regression assumes that the outcome variable comes from a bernoulli distribution (which is a special case of binomial distributions) where the number of trial $n$ is 1 and thus the outcome variable can only be 1 or 0. In contrast, binomial logistic regression assumes that the number of the target events follows a binomial distribution with $n$ trials and probability $q$. In this way, binomial logistic regression allows the outcome variable to take any non-negative integer value and thus is capable of handling count data. 

The Thai Educational Data records information about individual pupils that are clustered within schools. By aggregating the number of pupils who repeated a grade by school, we obtain a new data set where each row represents a school, with information about the proportion of pupils repeating a grade in that school. The `MSESC` (mean SES score) is also on the school level; therefore, it can be used to predict proportion or count of pupils who repeat a grade in a particular school. See below.

#### 6.1. Tranform Data
```{r}
ThaiEdu_Prop <- ThaiEdu_New %>%
  group_by(SCHOOLID, MSESC) %>%
  summarise(REPEAT = sum(REPEAT),
            TOTAL = n()) %>%
  ungroup()

head(ThaiEdu_Prop)
```

In this new data set, `REPEAT` refers to the number of pupils who repeated a grade; `TOTAL` refers to the total number of students in a particular school. 

#### 6.2. Explore data
```{r}
ThaiEdu_Prop %>%
  ggplot(aes(x = exp(MSESC)/(1+exp(MSESC)), y = REPEAT/TOTAL)) +
  geom_point() +
  geom_smooth(method = "lm")
```

We can see that the proportion of students who repeated a grade is negatively related to the inverse-logit of `MSESC`. Note that we model the variable `MSESC` as its inverse-logit because in a binomial regression model, we assume a linear relationship between the inverse-logit of the linear predictor and the outcome (i.e. proportion of events), not linearity between the predictor itself and the outcome.

#### 6.3. Fit a Binomial Logistic Regression Model
To fit a binomial logistic regression model, we also use the `glm` function. The only difference is in the specification of the outcome variable in the formula. We need to specify both the number of target events (`REPEAT`) and the number of non-events (`TOTAL-REPEAT`) and wrap them in `cbind()`.

```{r}
Model_Prop <- glm(formula = cbind(REPEAT, TOTAL-REPEAT) ~ MSESC,
                  family = binomial(logit),
                  data = ThaiEdu_Prop)

summary(Model_Prop)
```

#### 6.4. Interpretation
The parameter interpretation in a binomial regression model is the same as that in a binary logistic regression model. We know from the model summary above that the mean SES score of a school is negatively related to the odds of students repeating a grade in that school. To enhance interpretability, we use the `summ()` function again to calculate the exponentiated coefficient estimate of `MSESC`. Since `MSESC` is a continous variable, we can standardise the exponentiated `MSESC` estimate (by multiplying the original estimate with the SD of the variable, and then then exponentiating the resulting number).

```{r}
#Note that to use the summ() function for a binomial regression model, we need to make the outcome variable explicit objects:
REPEAT <- pull(filter(ThaiEdu_Prop, !is.na(MSESC)), REPEAT)
TOTAL <- pull(filter(ThaiEdu_Prop, !is.na(MSESC)), TOTAL)
summ(Model_Prop, exp = T, scale = T)
```

We can see that with a SD increase in `MSESC`, the odds of students repeating a grade is lowered by 1 - 85% = 15%.


We can visualise the effect of `MSESC`. 
```{r}
plot(allEffects(Model_Prop))
```


The plot above shows the expected influence of `MSESC` on the probability of a pupil repeating a grade. Holding everything else constant, as `MSESC` increases, the probability of a pupil repeating a grade lowers (from 0.19 to 0.10). The blue shaded areas indicate the 95% confidence intervals of the predicted values at each value of `MSESC`.


## **7. Multilevel Binary Logistic Regression**
The binary logistic regression model introduced earlier is limited to modelling the effects of pupil-level predictors; the binomial logistic regression is limited to modelling the effects of school-level predictors. To incorporate both pupil-level and school-level predictors, we can use multilevel models, specifically, multilevel binary logistic regression. If you are unfamiliar with multilevel models, you can use [Multilevel analysis: Techniques and applications](https://www.rensvandeschoot.com/multilevel-book/) for reference and [this tutorial](https://www.rensvandeschoot.com/tutorials/lme4/) for a good introduction to multilevel models with the `lme4` package in R. 


In addition to the motivation above, there are more reasons to use multilevel models. For instance, as the data are clustered within schools, it is likely that pupils from the same school are more similar to each other than those from other schools. Because of this, in one school, the probability of a pupil repeating a grade may be high, while in another school, low. Furthermore, even the relationship between the outcome (i.e. repeating a grade) and the predictor variabales (e.g. gender, preschool education, SES) may be different across schools. Also note that there are missing values in the `MSESC` variable. Using multilevel models can appropriately address these issues. 


See the following plot as an example. The plot shows the proportions of students repeating a grade across schools. We can see vast differences across schools. Therefore, we may need multilevel models.
```{r}
ThaiEdu_New %>%
  group_by(SCHOOLID) %>%
  summarise(PROP = sum(REPEAT)/n()) %>%
  plot()
```

We can also plot the relationship between `SEX` and `REPEAT` by `SCHOOLID`, to see whether the relationship between gender and repeating a grade differs by school.
```{r}
ThaiEdu_New %>%
  mutate(SEX = if_else(SEX == "boy", 1, 0)) %>%
  ggplot(aes(x = SEX, y = REPEAT, color = as.factor(SCHOOLID))) +
  geom_point(alpha = .1, position = "jitter")+
  geom_smooth(method = "glm", se = F, 
              method.args = list(family = "binomial")) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(0, 1)) +
  scale_y_continuous(breaks = c(0, 1))
```


In the plot above, different colors represent different schools. We can see that the relationship between `SEX` and `REPEAT` appears to be quite different across schools.


We can make the same plot for `PPED` and `REPEAT`.
```{r}
ThaiEdu_New %>%
  mutate(PPED = if_else(PPED == "yes", 1, 0)) %>%
  ggplot(aes(x = PPED, y = REPEAT, color = as.factor(SCHOOLID))) +
  geom_point(alpha = .1, position = "jitter")+
  geom_smooth(method = "glm", se = F, 
              method.args = list(family = "binomial")) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(0, 1)) +
  scale_y_continuous(breaks = c(0, 1))
```


The relationship between `PPED` and `REPEAT` also appears to be quite different across schools. However, we can also see that most of the relationships follow a downward trend, going from 0 (no previous schooling) to 1 (with previous schooling), indicating a negative relationship between `PPED` and `REPEAT`. 


Because of the observations above, we can conclude that there is a need for multilevel modelling in the current data, with not only a random intercept (`SCHOOLID`) but potentially also random slopes of the `SEX` and `PPED`. 


#### 7.1. Center Variables
Prior to fitting a multilevel model, it is necessary to center the predictors by using an appropriately chosen centering method (i.e. grand-mean centering or within-cluster centering), because the centering approach matters for the interpretation of the model estimates. Following the advice of Enders and Tofighi (2007), we should use within-cluster centering for the first-level predictors `SEX` and `PPED`, and grand-mean centering for the second-level predictor `MSESC`. 

```{r}
ThaiEdu_Center <- ThaiEdu_New %>%
  mutate(SEX = if_else(SEX == "girl", 0, 1),
         PPED = if_else(PPED == "yes", 1, 0)) %>%
  group_by(SCHOOLID) %>%
  mutate(SEX = SEX - mean(SEX),
         PPED = PPED - mean(PPED)) %>%
  ungroup() %>%
  mutate(MSESC = MSESC - mean(MSESC, na.rm = T))

head(ThaiEdu_Center)
```


#### 7.2. Intercept Only Model
To specify a multilevel model, we use the `glmer` function from the `lme4` package. Note that the random effect term should be included in parentheses. In addition, within the parentheses, the random slope term(s) and the cluster terms should be separated by `|`. Note that we use an additional argument `control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))` in the `glmer` function to specify a higher number of maximum iterations than default (10000). This might be necessary because a multilevel model may require a large number of iterations to converge.  

  
We start by specifying an intercept-only model, in order to assess the impact of the clustering structure of the data. 
```{r}
Model_Multi_Intercept <- glmer(formula = REPEAT ~ 1 + (1|SCHOOLID),
                               family = binomial(logit),
                               data = ThaiEdu_Center,
                               control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)))

summary(Model_Multi_Intercept)
```

Below we calculate the ICC (intra-class correlation) of the intercept-only model.
```{r}
icc(Model_Multi_Intercept)
```

An ICC of 0.33 means that 33% of the variation in the outcome variable can be accounted for by the clustering stucture of the data. This provides evidence that a multilevel model may make a difference to the model estimates, in comparison with a non-multilevel model. Therefore, the use of multilevel models is necessary and warrantied.  

#### 7.3. Full Model
It is good practice to build a multilevel model step by step. However, as this tutorial's focus is not on muitilevel modelling, we go directly from the intercept-only model to the full-model that we are ultimately interested in. In the full model, we include not only fixed effect terms of `SEX`, `PPED` and `MSESC` and a random intercept term, but also random slope terms for `SEX` and `PPED`. Note that we specify `family = binomial(link = "logit")`, as this model is essentially a binary logistic regression model. 

```{r}
Model_Multi_Full <- glmer(REPEAT ~ SEX + PPED + MSESC + (1 + SEX + PPED|SCHOOLID),
                          family = binomial(logit),
                          data = ThaiEdu_Center,
                          control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)))
summary(Model_Multi_Full)
```

The results (pertaining to the fixed effects) are similar to the results of the previous binary logistic regression and binomial logistic regression models. On the pupil-level, `SEX` has a significant and positive influence on the odds of a pupil repeating a grade, while `PPED` has a significant and negative influence. On the school-level, `MSESC` has a significant and negative effect on the outcome variable. Let's also look at the variance of the random effect terms. 

Again, we can use the `summ()` function to retrieve the exponentiated coefficient estimates for easier interpretation. 
```{r}
summ(Model_Multi_Full, exp = T)
```

We can also use the `allEffects` function to visualise the effects of the parameter estimates. Note that because the first-level categorical variables (`SEX` and `PPED`) are centered, they are treated as continuous variables in the model and as well in the following effect plots. 
```{r}
plot(allEffects(Model_Multi_Full))
```

In addition to the fixed-effect terms, let's also look at the random effect terms. From the ICC value before, we know that it's necessary to include a random intercept. However, the necessity of including random slopes for `SEX` and `PPED` is less clear. To find this out, we can use the likelihood ratio test and AIC to judge whether the inclusion of the random slope(s) improves model fit.

```{r}
#let's fit a less-than-full model that leaves out the random slope term of `SEX`
Model_Multi_Full_No_SEX <- glmer(REPEAT ~ SEX + PPED + MSESC + (1 + PPED|SCHOOLID),
                                 family = binomial(logit),
                                 data = ThaiEdu_Center,
                                 control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)))

#let's fit a less-than-full model that leaves out the random slope term of `PPED`
Model_Multi_Full_No_PPED <- glmer(REPEAT ~ SEX + PPED + MSESC + (1 + SEX|SCHOOLID),
                                  family = binomial(logit),
                                  data = ThaiEdu_Center,
                                  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)))

#let's fit a less-than-full model that leaves out the random slope terms of both `SEX` and `PPED`
Model_Multi_Full_No_Random_Slope <- glmer(REPEAT ~ SEX + PPED + MSESC + (1|SCHOOLID),
                                      family = binomial(logit),
                                      data = ThaiEdu_Center,
                                      control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)))

```

Likelihood ratio test:
```{r}
#compare the full model with that model that excludes `SEX`
anova(Model_Multi_Full_No_SEX, Model_Multi_Full, test="Chisq")
```

```{r}
#compare the full model with that model that excludes `PPED`
anova(Model_Multi_Full_No_PPED, Model_Multi_Full, test="Chisq")
```

```{r}
anova(Model_Multi_Full_No_Random_Slope, Model_Multi_Full, test="Chisq")
```

From the all insignificant likelihood ratio test results (Pr(>Chisq) > 0.05), we can conclude that there is no significant improvement in model fit by adding any random slope terms. 


AIC:
```{r}
AIC(logLik(Model_Multi_Full)) #full model
AIC(logLik(Model_Multi_Full_No_SEX)) #model without SEX
AIC(logLik(Model_Multi_Full_No_PPED)) #model without PPED
AIC(logLik(Model_Multi_Full_No_Random_Slope)) #model without random slopes
```
From the AIC results, we see that including random slope terms either does not substantially improve AIC (indicated by lower AIC value) or leads to worse AIC (i.e. higher). Therefore, we also conclude there is no need to include the random effect term(s).

## **8. Other Family (Distribution) and Link Functions**
So far, we have introduced binary and binomial logistic regression, both of which come from the binomial family with the logit link. However, there are many more distribution families and link functions that we can use in glm analysis. For instance, to model binary outcomes, we can also use the probit link or the complementary log-log (cloglog) instead of the logit link. To model count data, we can also use Poisson regression, which assumes that the outcome variable comes from a Poisson distribution and uses the logarithm as the link function. For an overview of possible glm models, see the [Wikipedia page](https://en.wikipedia.org/wiki/Poisson_regression) for GLM.



## **References**
Bates, D., Maechler, M., Bolker, B., & Walker, S. (2015). *Fitting Linear Mixed-Effects Models Using lme4. Journal of Statistical Software, 67*(1), 1-48. [doi:10.18637/jss.v067.i01](https://doi.org/10.18637/jss.v067.i01)


Enders, C. K., & Tofighi, D. (2007). Centering predictor variables in cross-sectional multilevel models: A new look at an old issue. *Psychological Methods, 12*(2), 121-138. [doi:10.1037/1082-989X.12.2.121](https://doi.org/10.1037/1082-989X.12.2.121)


Fox, J. (2003). *Effect Displays in R for Generalised Linear Models. Journal of Statistical Software, 8*(15), 1-27. [http://www.jstatsoft.org/v08/i15/](http://www.jstatsoft.org/v08/i15/) 


Long, JA. (2019). jtools: Analysis and Presentation of Social Scientific Data. R package version 2.0.1, [https://cran.r-project.org/package=jtools](https://cran.r-project.org/package=jtools)


L&uuml;decke, D. (2019). sjstats: Statistical Functions for Regression Models (Version 0.17.5). [doi: 10.5281/zenodo.1284472](http://dx.doi.org/10.5281/zenodo.1284472)


Raudenbush, S. W., & Bhumirat, C. (1992). The distribution of resources for primary education and its consequences for educational achievement in Thailand. *International Journal of Educational Research, 17*(2), 143-164. [doi:10.1016/0883-0355(92)90005-Q](https://doi.org/10.1016/0883-0355(92)90005-Q)


Sing, T., Sander, O., Beerenwinkel, N. & Lengauer, T. (2005). *ROCR: visualizing classifier performance in R. Bioinformatics, 21*(20), pp. 7881. [http://rocr.bioinf.mpi-sb.mpg.de](http://rocr.bioinf.mpi-sb.mpg.de)


Wickham, H. (2017). tidyverse: Easily Install and Load the 'Tidyverse'. R package version 1.2.1. [https://CRAN.R-project.org/package=tidyverse](https://CRAN.R-project.org/package=tidyverse)